Tree size, microhabitat, and hydraulic traits shape drought responses in a temperate broadleaf forest

### 2019-09-20 manuscript/climate_sensitivity_draft.Rmd
worked on abstract and discussion

### Summary
Predicting forest responses to drought is an increasingly critical task under climate change effects.
Part of the problem is due to the lack of studies analyzing the confluence of leaf hydraulic traits with biophysical parameters.
In this study, we analyze the interaction between these two trait groups using forest census data from a 25.6-ha ForestGEO plot in Virginia (USA).
Drought periods were defined by both Palmer Drought Severity Indices (PDSI) and their identification from tree-ring records for 12 species representing 97% of woody productivity.
Each drought scenario (1966, 1977, 1999), along with the overall trend, was then tested against leaf hyraulic trait measurements and microhabitat biophysical data.
Individual-level growth responses to the three individual droughts  were stronger in three cases: taller trees in dominant canopy positions, trees in wetter microsites, and more drought-sensitive species as assessed by leaf traits (turgor loss at less negative leaf water potential, greater shrinkage with leaf dehydration). However, there was substantial variation in the best predictor variables across given droughts.
We conclude that when droughts occur, large dominant trees, drought-sensitive species, and individuals in wetter microhabitats are likely to be most strongly affected. **Add discussion points**

@@

### Introduction
Forests globally play a critical role in climate regulation ([Bonan 2008](https://science.sciencemag.org/content/320/5882/1444)), yet there remains enormous uncertainty as to how the terrestrial carbon (C) sink, which is dominated by forests, will respond to climate change ([Friedlingstein et al. 2006](https://journals.ametsoc.org/doi/full/10.1175/JCLI3800.1)). An important aspect of this uncertainty lies in responses to drought (**REF**). In many forested regions around the world, the risk of severe drought is increasing ([Trenbert et al. 2014](https://www.nature.com/articles/nclimate2067)), even in conjunction with increasing precipitation ([IPCC 2014](https://www.cambridge.org/core/books/climate-change-2014-impacts-adaptation-and-vulnerability-part-b-regional-aspects/036A899BD52861D61B0D519C5F2B9334)). Global change-type drought has been affecting forests worldwide ([Allen et al. 2015](https://doi.org/10.1016/j.foreco.2009.09.001)), and it is expected that future climate change-driven droughts will severely impact forests around the world ([Allen et al. 2015](https://doi.org/10.1016/j.foreco.2009.09.001); **REFS**). Larger trees tend to suffer more (e.g., [@bennett_larger_2015]; Stovall et al. 2019), resulting in disproportionate impacts on forest C storage ([Meakem et al. 2018](https://nph.onlinelibrary.wiley.com/doi/10.1111/nph.14633)). As a result, forest drought responses stand to strongly impact forest feedbacks to climate change (**REFS**), yet accurate characterization of drought responses remains a modeling challenge (**REFS**)-- in part because some of the mechanisms underlying drought responses remain unclear. Understanding forest responses to drought requires increased functional understanding of how tree size, microhabitat, and species' traits jointly confer individual-level vulnerability or resistance, and the extent to which their influence is consistent across droughts.

@@

\raggedright

\newpage
### Results

@@

\newpage
### Discussion
Discussion outline

>>>>

### Summary (**needs work**)
- Predicting forest responses to drought is an increasingly critical task under climate change effects.
[Part of the problem is due to the lack of studies analyzing the confluence of leaf hydraulic traits with biophysical parameters.]

- In this study, we analyze tree-ring records for 12 species representing 97% of woody productivity in the 25.6-ha ForestGEO plot in Virginia (USA) to determine how tree size, microhabitat, and species' traits interactively shape drought responses across the three strongest droughts over the 60 year period from 1950 and 2009.

- Individual-level growth responses to the three individual droughts  were stronger in three cases: taller trees in dominant canopy positions, trees in wetter microsites, and more drought-sensitive species as assessed by leaf traits (turgor loss at less negative leaf water potential, greater shrinkage with leaf dehydration). However, there was substantial variation in the best predictor variables across given droughts.

- We conclude that when droughts occur, large dominant trees, drought-sensitive species, and individuals in wetter microhabitats tend to be most strongly affected.

@@

\newpage
### Introduction
Forests globally play a critical role in climate regulation ([Bonan 2008](https://science.sciencemag.org/content/320/5882/1444)), yet there remains enormous uncertainty as to how the terrestrial carbon (C) sink, which is dominated by forests, will respond to climate change ([Friedlingstein et al. 2006](https://journals.ametsoc.org/doi/full/10.1175/JCLI3800.1)). An important aspect of this uncertainty lies in responses to drought (**REF**). In many forested regions around the world, the risk of severe drought is increasing ([Trenbert et al. 2014](https://www.nature.com/articles/nclimate2067)), even in conjunction with increasing precipitation ([IPCC 2014](https://www.cambridge.org/core/books/climate-change-2014-impacts-adaptation-and-vulnerability-part-b-regional-aspects/036A899BD52861D61B0D519C5F2B9334)). Global change-type drought has been affecting forests worldwide ([Allen et al. 2015](https://doi.org/10.1016/j.foreco.2009.09.001)), and it is expected that future climate change-driven droughts will severely impact forests around the world ([Allen et al. 2015](https://doi.org/10.1016/j.foreco.2009.09.001); **REFS**). Larger trees tend to suffer more (e.g., [@bennett_larger_2015]; Stovall et al. 2019), resulting in disproportionate impacts on forest C storage ([Meakem et al. 2018](https://nph.onlinelibrary.wiley.com/doi/10.1111/nph.14633)). As a result, forest drought responses stand to strongly impact forest feedbacks to climate change (**REFS**), yet accurate characterization of drought responses remains a modeling challenge (**REFS**)-- in part because some of the mechanisms underlying drought responses remain unclear. Understanding forest responses to drought requires increased functional understanding of how tree size, microhabitat, and species' traits jointly confer individual-level vulnerability or resistance, and the extent to which their influence is consistent across droughts.

@@

### Results

@@

### Discussion

Our results reveal how tree size, microhabitat, and hydraulic traits shape tree growth responses across three droughts in a temperate deciduous forest. The tendencey for larger trees to suffer more was driven primarily by their height, with a marginal effect of crown exposure.

\newpage
Discussion outline
